Temperature Effects on Surface Sulfur Dioxide

### 2011-08-24 Cycle-Results.tex
Almost done adding content to the cycle results. Only waiting to write up the energy results.

Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule. The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier and depicted in Figure \ref{fig:cyclic-structures}. Only the most commonly occuring subset of the cyclic structures were analyzed.

@@

Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic structure? To answer the questions of cyclic lifespans, a method was devised to define a lifetime of a cycle. For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was found or not found, respectively. This resulted in a time-function, $C(t)$, similar in nature to a digital signal. The resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the signal is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed. The cyclic structure is very dynamic constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the signal lasting less than 10 fs each. Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to ``debounce'' the time-signal to disregard very brief moments when the cycle breaks or forms for less than 20 fs. The time-signal, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width. The resulting smoothed signal, $C_s(t)$, was then cutoff with the following criteria:

@@

@@

>>>>

Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule. The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier and depicted in Figure \ref{fig:cyclic-structures}. Only the most commonly occurring subset of the cyclic structures were analyzed.

@@

Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic structure? To answer the questions of cyclic lifespans, a method was devised to define a lifetime of a cycle. For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was found or not found, respectively. This resulted in a time-function, $C(t)$, similar in nature to a digital signal. The resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed. The cyclic structure is very dynamic constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the function lasting less than 10 fs each. Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to ``debounce'' the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs. The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

@@

where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid ``bouncing'' of the cyclic time-function, $C(t)$. Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final ``debounced'' function, $f(t)$ (green).

@@

The distribution of cycle lifetimes (i.e. contiguous spans of time spent with $f(t) = 1$) was determined. A plot of the cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations. The most frequent lifespan for both temperature data sets lasts between 0-1 ps. This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time. Even with the ``debouncing'' procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps. The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend also extending to the longer lifespans as well. The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps. All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes. Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time.

Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles. Above 4 ps, most of the cycles that form are found in the hot system. The 8 ps cycles are notable in that they form  for just under half of an entire single simulated trajectory.

We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently. The \suldiox~is likely forming and breaking bonds with waters extraneous to the cyclic hydrate structures (i.e. not directly involved in the bonds of the cycle). For the longer-lived cycles, the extraneous bonding to the \suldiox~may have little effect on the cyclic hydrate waters. However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e. ``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken. Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure. Figure \ref{fig:bonding-coordinations} shows the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively. Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold). The increased temperature appears to cause the \suldiox~to bond in such a way that is more conducive to the formation of cyclic structures. Furthermore, the distribution of bonding coordinations in the hot system shifts the transition between bonding coordinations away from cycle-breaking coordinations (i.e. those coordinations that necessarily disallow cycle formation). This possibly accounts for the slightly increased populations of longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/cycles/cyclic-lifespans-inset-small.png}
		\caption{}
		\label{fig:cycle-lifespans}
	\end{center}
\end{figure}
